---
title: "This is a reproducible R Markdown doc"
author: "Weijia Chen"
date: "3 September 2014"
output: html_document
---

###### Step one: Merges the training and the test sets to create one data set, the final mergedfull object realise the prototype table data frame. See the prototype table go to the README.md file. Look at the output: 10299 * 563

```{r, echo=FALSE}
setwd("~/Getting-and-Cleaning-Data/project.solution/") #set the working director at "~/Users/user's homefolder/Getting-and-Cleaning-Data/project.solution/"
library(dplyr) #load library dplyr and it will be used to manipulate data
traindata <- read.table("./project.data/train/X_train.txt") #load data from X_train.txt
traindata <- tbl_df(traindata) #tbl objects only print a few rows and all the columns that fit on one screen
#str(traindata) #checck traindata dimensions. 7352 (rows) * 561 (cols); visualise the table and get the column names from V1 to V561
testdata<- read.table("./project.data/test/X_test.txt") #load data from X_test.txt
testdata <- tbl_df(testdata); #str(testdata) #check testdata dimensions. 2947 (rows) * 561(cols); The final data frame should have (7352 + 2947) rows; traindata accounting for around 70% of total data and testdata accounting for around 30% of total data
mergeddata <- tbl_df(rbind_list(traindata,testdata)); #str(mergeddata) #10299 (rows) * 561 (cols), with cols name from V1 to V561

features <- read.table("./project.data/features.txt", stringsAsFactors = FALSE) #load variable names from features.txt. Second col should be character class, thus, second statement is stringsAsFactors = FALSE

colnames(mergeddata) <- features$V2; #str(mergeddata) #change megeddata variable names from (V1 - V561) to tBodyAcc-mean()-X...tBodyAccJerk-energy()-Z 

trainacts <- read.table("./project.data/train/y_train.txt") #load activities data from y_train.txt
trainacts <- tbl_df(trainacts); #str(trainacts) #check trainacts dimensions. 7352 (rows) * 1 (cols), col name is V1
testacts <- read.table("./project.data/test/y_test.txt") #load activities data from y_test.txt
testacts <- tbl_df(testacts); #str(testacts) #check testacts dimensions. 2947 (rows) * 1 (cols), col name is V1
mergedacts <- tbl_df(rbind_list(trainacts, testacts)); #str(mergedacts) #check mergedacts dimensions. 10299 (rows) * 1 (col), with col name V1
mergedacts <- select(mergedacts, activities = V1); #str(mergedacts) #change variable name V1 to activities, in order to distinguish with the V1 variable of megeddata

trainsubject <- read.table("./project.data/train/subject_train.txt") #load subject data from subject_train.txt
trainsubject <- tbl_df(trainsubject); #str(trainsubject) #check trainsubject dimensions. 7352 (rows) * 1 (cols, V1) 
testsubject <- read.table("./project.data/test/subject_test.txt") #load subject data from subject_test.txt
testsubject <- tbl_df(testsubject); #str(testsubject) #check testsubject dimensions. 2947 (rows) * 1 (cols, V1)
mergedsubject <- tbl_df(rbind_list(trainsubject, testsubject)); #str(mergedsubject) #check mergedsubject dimensions. 10299 (rows) * 1 (cols, V1)
mergedsubject <- select(mergedsubject, volunteer = V1); #str(mergedsubject) #change variable name V1 to volunteer, in order to distinguish with the V1 variable of megeddata

mergedfull <- cbind(mergedsubject, mergedacts, mergeddata) #combine three data frame into one with 10299 (rows) * 563 (cols)
mergedfull <- tbl_df(mergedfull); str(mergedfull) #check mergedfull dimensions. 
```

###### Step two: Extracts only the measurements on the mean and standard deviation for each measurement. Look at the output: 10299 * 88, activities: int

```{r, echo=FALSE}
mergedfullsubset <- select(mergedfull, volunteer, activities, contains("mean"), contains("std")); str(mergedfullsubset)
```

##### Step three: Uses descriptive activity names to name the activities in the data set. Look at the output: 10299 * 88, activity: Factor

```{r, echo=FALSE}
act.labels <- read.table("./project.data/activity_labels.txt")
names(act.labels) <- c("activities", "activity") 
mergedfullsubset <- inner_join(mergedfullsubset, act.labels); #str(mergedfullsubset); 
mergedfullsubset <- select(mergedfullsubset, volunteer, activity, contains("mean"),contains("std")); str(mergedfullsubset) 
```


##### Step five: no output of step four. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject. Look at the output: 10299 * 88

```{r, echo=FALSE}
tidydata <- read.table("cleaneddata.txt") 
tidydata <- tbl_df(tidydata); #str(tidydata) 
tidydatamean <- tidydata %>% 
        group_by(volunteer, activity) %>% 
        summarise_each(funs(mean(., na.rm = TRUE)), matches("mean"),matches("std")) 

tidydatamean
write.table(tidydatamean,"tidydatamean.txt") 
write.csv(tidydatamean, "tidydatamean.csv") 
```

# © Weijia Chen Reserves All Rights, 2014.
